On the properties of discs around accreting brown dwarfs

### 2010-07-01 BDD_submit/paper.tex
Changed cuts, Taurus cut was applied by monin_2010 but was from gutermuth_2008
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2340 25464f21-6099-402c-b30a-640f1d40d38b

In surveys, particularly those attempting to discern the PMS disc
fraction this disentangling usually takes place using broad-band
photometric measures and cuts in colour-colour or colour-magnitude
space \citep{luhman_2005b,luhman_2008,gutermuth_2008}.  The important
quantity here is the separation in wavelength between the emission
peaks for the stellar and thermal disc components, as it is these
components which must be isolated. For BD disc (BDD hereafter) systems
it is likely that difficulties detecting the disc will be exacerbated
by the lower temperatures of the photosphere and therefore the smaller
separation in wavelength from the thermal disc emission component,
compared to CTTS systems. Therefore, to derive disc fractions for BD
populations we require detailed comparison models with which to guide
disc candidate selection.

@@

Here we present a grid of models of BDD systems, including a
self-consistent treatment of the photospheric and accretion luminosity
sources and the interaction of that flux with the circumstellar disc.
We investigate the impact of the luminosity of the central source on
location and shape of the disc inner rim, as well as the large-scale
structure and flaring of the outer disc. We construct synthetic
colour-colour and colour-magnitude diagrams in order to examine the
efficacy of the photometric selections used to isolate brown dwarfs
and measure disc fractions
\citep{luhman_2005b,luhman_2008,gutermuth_2008}.

@@

Disc fractions have been derived using infrared excesses previously in
\textit{JHK} \citep[e.g][]{lada_1995,carpenter_1997}, however recent
works pre-dominantly use \textit{Spitzer} IRAC magnitudes
\citep[e.g][]{luhman_2005b,luhman_2008,gutermuth_2008}. Furthermore,
MIPS magnitudes are used to identify so-called debris discs, where IR
excesses are not apparent at shorter wavelengths
\citep[e.g][]{currie_2008,bryden_2009}. Finally, disc fractions have
also been derived using the $\alpha$ criteria, where
$\alpha=\frac{dlog\lambda F\lambda}{dlogF\lambda}$ between two
limiting wavelengths, originally used to distinguish amongst Class I,
II or II sources, but now used to detect disc presence
\citep{lada_2006,kennedy_2009}. An $\alpha>-2$ is used as a selection
criterion for disc presence for TTS stars. We have constructed the
$\alpha$ values for our model grid by adopting the limiting
wavelengths of \cite{kennedy_2009}, namely 3.6 to 8.0$\mu$m.

@@

The observational cuts applied, shown as dashed lines, are from
\cite{luhman_2005b} a study of IC348, \cite{luhman_2008} a study of
$\sigma$ Orionis and \cite{gutermuth_2008} a study of the Taurus
region. In these cases the effects of extinction are either negligible
in the plotted colours, with values of $E([3.6]-[4.5])<0.04$ and
$E([4.5]-[5.8])<0.02$ for IC348 and $A_V \leq$4 mag \citep[which will
be negligible in the IRAC CoCoD,][]{allen_2004}, or the cuts
have been placed in intrinsic colour space as for $\sigma$
Orionis. The cuts from \cite{luhman_2005b}, \cite{luhman_2008} and
\cite{gutermuth_2008} appear as the top, top-middle and bottom-middle
panels. The lower panel shows a recent BDD candidate selection using
the $\alpha$ value from \citep{kennedy_2009}.

\begin{figure*}
  \centering
  \includegraphics[scale=0.65,angle=90]{./Fig/disc_cuts.ps}
  \caption{Figure showing both from top to bottom, the disc candidate
    selection of \citet{luhman_2005b}, \citet{luhman_2008},
    \citet{gutermuth_2008} and \citet{kennedy_2009}. The photometry or
    $\alpha$ values for all models are shown, with naked stars as
    black crosses, 1 Myr BDD systems as blue dots and 10 Myr BDD
    systems as red dots. The dashed horizontal and vertical lines are
    disc selections from the studies in question. The left panels are
    the typical accretors and the right panels the extreme
    accretors. The top three panels are IRAC CoCoDs and the bottom
    panel plots the $\alpha$ value against run number (an arbitary
    number).\label{disc_cuts}}
\end{figure*}

Figure \ref{disc_cuts} shows that the recent observational cuts used
to identify disc candidates would be, in the main, reliable for our
model grid. Simulated photometry would be correctly identified using
these cuts. It is important to note that our conclusions so far have
been drawn from differential photometric arguments, in this case we
are using intrinsic colours and these values are extremely sensitive
to changes in zero point and photometric calibration. The top panel of
Figure \ref{disc_cuts} shows that other than a few typically accreting
systems the observational cut of \cite{luhman_2005b} would select all
of the BDD systems in our grid. The second panel down shows that the
cut of \cite{luhman_2008} would miss some typical and a very small
number of extreme accreting BDD systems (almost all are edge-on
systems). The third panel down shows only very few BDD systems will be
missed by the selection of \cite{gutermuth_2008}. Finally, the
selection for CTTS candidates of $\alpha>-$2 \citep{kennedy_2009}
appears reasonably applicable to our BDD systems. As can be seen in
the bottom panel of Figure \ref{disc_cuts} almost all the typical, and
all of the extreme accreting BDD systems would be successfully
identified using this criterion. Again the missed typically accreting
systems are edge-on systems. Given, that extreme reddening may move
naked stars into the BDD region selected, in general, disc fractions
are usually quoted as lower limits. Therefore, a small number of
miss-identified BDD system is a negligible effect. This allows us to
conclude, that ubiquitously used BDD selections within IRAC CoCoDs are
reliable when applied to our model grid. Additionally, the $\alpha$
value would be a reliable disc indicator for our model grid.

>>>>

In surveys, particularly those attempting to discern the PMS disc
fraction this disentangling usually takes place using broad-band
photometric measures and cuts in colour-colour or colour-magnitude
space \citep{luhman_2005b,luhman_2008,gutermuth_2008,monin_2010}.  The
important quantity here is the separation in wavelength between the
emission peaks for the stellar and thermal disc components, as it is
these components which must be isolated. For BD disc (BDD hereafter)
systems it is likely that difficulties detecting the disc will be
exacerbated by the lower temperatures of the photosphere and therefore
the smaller separation in wavelength from the thermal disc emission
component, compared to CTTS systems. Therefore, to derive disc
fractions for BD populations we require detailed comparison models
with which to guide disc candidate selection.

@@

Here we present a grid of models of BDD systems, including a
self-consistent treatment of the photospheric and accretion luminosity
sources and the interaction of that flux with the circumstellar disc.
We investigate the impact of the luminosity of the central source on
location and shape of the disc inner rim, as well as the large-scale
structure and flaring of the outer disc. We construct synthetic
colour-colour and colour-magnitude diagrams in order to examine the
efficacy of the photometric selections used to isolate brown dwarfs
and measure disc fractions
\citep{luhman_2005b,luhman_2008,gutermuth_2008,monin_2010}.

@@

Disc fractions have been derived using infrared excesses previously in
\textit{JHK} \citep[e.g][]{lada_1995,carpenter_1997}, however recent
works pre-dominantly use \textit{Spitzer} IRAC magnitudes
\citep[e.g][]{luhman_2005b,luhman_2008,gutermuth_2008,monin_2010}. Furthermore,
MIPS magnitudes are used to identify so-called debris discs, where IR
excesses are not apparent at shorter wavelengths
\citep[e.g][]{currie_2008,bryden_2009}. Finally, disc fractions have
also been derived using the $\alpha$ criteria, where
$\alpha=\frac{dlog\lambda F\lambda}{dlogF\lambda}$ between two
limiting wavelengths, originally used to distinguish amongst Class I,
II or II sources, but now used to detect disc presence
\citep{lada_2006,kennedy_2009}. An $\alpha>-2$ is used as a selection
criterion for disc presence for TTS stars. We have constructed the
$\alpha$ values for our model grid by adopting the limiting
wavelengths of \cite{kennedy_2009}, namely 3.6 to 8.0$\mu$m.

@@

The observational cuts applied, shown as dashed lines, are from
\cite{luhman_2005b} a study of IC348, \cite{luhman_2008} a study of
$\sigma$ Orionis and \cite{monin_2010} \citep[using the criteria
of][]{gutermuth_2008} a study of the Taurus region. In these cases the
effects of extinction are either negligible in the plotted colours,
with values of $E([3.6]-[4.5])<0.04$ and $E([4.5]-[5.8])<0.02$ for
IC348 and $A_V \leq$4 mag \citep[which will be negligible in the IRAC
CoCoD,][]{allen_2004}, or the cuts have been placed in intrinsic
colour space as for $\sigma$ Orionis. The cuts from
\cite{luhman_2005b}, \cite{luhman_2008} and \cite{monin_2010} appear
as the top, top-middle and bottom-middle panels. The lower panel shows
a recent BDD candidate selection using the $\alpha$ value from
\citep{kennedy_2009}.

\begin{figure*}
  \centering
  \includegraphics[scale=0.65,angle=90]{./Fig/disc_cuts.ps}
  \caption{Figure showing both from top to bottom, the disc candidate
    selection of \citet{luhman_2005b}, \citet{luhman_2008},
    \citet{monin_2010} \citep[using the criteria of][]{gutermuth_2008}
    and \citet{kennedy_2009}. The photometry or $\alpha$ values for
    all models are shown, with naked stars as black crosses, 1 Myr BDD
    systems as blue dots and 10 Myr BDD systems as red dots. The
    dashed horizontal and vertical lines are disc selections from the
    studies in question. The left panels are the typical accretors and
    the right panels the extreme accretors. The top three panels are
    IRAC CoCoDs and the bottom panel plots the $\alpha$ value against
    run number (an arbitary number).\label{disc_cuts}}
\end{figure*}

Figure \ref{disc_cuts} shows that the recent observational cuts used
to identify disc candidates would be, in the main, reliable for our
model grid. Simulated photometry would be correctly identified using
these cuts. It is important to note that our conclusions so far have
been drawn from differential photometric arguments, in this case we
are using intrinsic colours and these values are extremely sensitive
to changes in zero point and photometric calibration. The top panel of
Figure \ref{disc_cuts} shows that other than a few typically accreting
systems the observational cut of \cite{luhman_2005b} would select all
of the BDD systems in our grid. The second panel down shows that the
cut of \cite{luhman_2008} would miss some typical and a very small
number of extreme accreting BDD systems (almost all are edge-on
systems). The third panel down shows only very few BDD systems will be
missed by the selection of \cite{gutermuth_2008} which was applied to
Taurus by\cite{monin_2010}. Finally, the selection for CTTS candidates
of $\alpha>-$2 \citep{kennedy_2009} appears reasonably applicable to
our BDD systems. As can be seen in the bottom panel of Figure
\ref{disc_cuts} almost all the typical, and all of the extreme
accreting BDD systems would be successfully identified using this
criterion. Again the missed typically accreting systems are edge-on
systems. Given, that extreme reddening may move naked stars into the
BDD region selected, in general, disc fractions are usually quoted as
lower limits. Therefore, a small number of miss-identified BDD system
is a negligible effect. This allows us to conclude, that ubiquitously
used BDD selections within IRAC CoCoDs are reliable when applied to
our model grid. Additionally, the $\alpha$ value would be a reliable
disc indicator for our model grid.
